### 2017-10-02 papers/Physica2016/biblio.tex

\section*{References}
\begin{thebibliography}{20}
\bibitem{Bell1978} Bell A R 1978 \textit{MNRAS} \textbf{182} 147
\bibitem{Blandford1978} Blandford R D and Ostriker J P 1978 \textit{Astrophys. J.} \textbf{221} L29 
\bibitem{Berezhko2003} Berezhko E G, Ksenofontov L T and V{\"o}lk H J  2003 \textit{A}{\&}\textit{A} \textbf{412} L11
\bibitem{Uchiyama2007} Uchiyama Y, Aharonian F A, Tanaka T, Takahashi T and Maeda Y 2007 \textit{Nature} \textbf{449} 576
\bibitem{Lapenta2006} Lapenta G, Brackbill J U and Ricci P 2006 \textit{Phys. Plasmas} \textbf{13} 055904
\bibitem{Noguchi2007} Noguchi K, Tronci C, Zuccaro G and Lapenta G 2007 \textit{Phys. Plasmas} \textbf{14} 042308
\bibitem{Weibel1959} Weibel E S 1959 \textit{Phys. Rev. Lett.} \textbf{2} 83
\bibitem{Yoon1987} Yoon P H and Davidson R C 1987 \textit{Phys. Rev. A} \textbf{35} 2718
\bibitem{Park2015} Park J, Caprioli D and Spitkovsky A 2015 \textit{Phys. Rev. Lett.} \textbf{114} 085003
\end{thebibliography}