### 2017-10-02 papers/Physica2016/intro.tex

\section{Introduction}
One of the possible sources of high-energy cosmic rays are shock waves in astrophysical collisionless plasma due to the diffusive shock acceleration (DSA) mechanism, see \cite{Bell1978}, \cite{Blandford1978}. Strong magnetic fields are necessary for effective acceleration of charged particles and the observations of the synchrotron emission from supernova remnants (SNRs) show, that the magnetic field there is about 100 times stronger then the interstellar field (see for example \cite{Berezhko2003},\cite{Uchiyama2007}). So strong fields can be produced due to instabilities in anisotropic plasma. One of the most powerful methods to explore processes in collisionless plasma is the Particle-in-Cell simulation. In this work we explore instabilities in the plasma using this method.