### 2017-11-10 papers/Physica2017/Romansky_Physica2017.tex
abstract

\documentclass[a4paper]{jpconf}
\bibliographystyle{iopart-num}
\usepackage{amsmath}
\usepackage{citesort}
\usepackage{subfigure}
\usepackage{graphicx}
\graphicspath{{fig/}}
\usepackage{ifpdf}
\ifpdf\usepackage{epstopdf}\fi
\usepackage[export]{adjustbox}

%----------------------------------------------------- 
%\usepackage{soul,ulem,color,xspace,bm}
% Andrei's commands
% Suggest to remove
%\newcommand{\asrm}[1]{{\color{magenta}\sout{#1}}}
% Suggest to insert
%\newcommand{\as}[1]{\color{cyan}#1\xspace\color{black}}
% Suggest to replace
%\newcommand{\asrp}[2]{\asrm{#1} \as{#2}}
% Comment
%\newcommand{\ascm}[1]{{\color{green}\;AS: #1}}
%------------------------------------------------------

\begin{document}
\title{Particle-in-cell simulation of ion and electron acceleration in trans-relativistic shocks.}

\author{V I Romansky$^{1,2}$, A M Bykov$^1$, S M Osipov$^1$ and P E Gladilin$^1$}

\address{$^1$ Ioffe Institute, 26 Politekhnicheskaya st., St. Petersburg 194021, Russia}
\address{$^2$ 
Sternberg Astronomical Institute, Moscow State University
 Universitetsky pr., 13, Moscow 119234, Russia}

\ead{romanskyvadim@gmail.com}

\begin{abstract}
The problem of Cosmic Ray (CR) origin and acceleration is a widely discussed problem of
modern astrophysics. It is considered that most likely CR sources are supernova remnant
shock waves. One of the methods to study particle injection and acceleration by shock
waves is a particle-in-cell computer simulation. In this paper we present results of shock
wave simulations based on our self-developed implicit particle-in-cell code. 
\end{abstract}
\input{intro.tex}
\input{model.tex}
\input{test.tex}
\input{shock.tex}
\input{conclusions.tex}
\input{biblio.tex}
\end{document}