### 2017-11-10 papers/Physica2017/conclusions.tex
fz

\section{Conclusions}
In this paper we presented the developed 3D implicit particle-in-cell code, which can be applied for various problems in the collisionless plasma. We have tested it and obtained results, that are consistent with theoretical predictions. We studied the influence of admixture of heavy ions on the evolution of the shock wave and simulations have shown that the presence of heavy ions increases the number of accelerated particles. Further research is necessary in order to study this phenomena in detail.
\ack
A M Bykov, P E  Gladilin and S M  Osipov have developed the implicit fields solver module, which provides conservation of the divergenceless magnetic field. 
V I Romansky acknowledges support from RSF grant 16-12-10519 which was used to develop the
particle mover, parallelization and perform the testing of the code.

The results of the work were obtained using computational resources of Peter the Great Saint-Petersburg Polytechnic University Supercomputing Center (http://www.spbstu.ru).