### 2017-11-11 papers/Physica2017/biblio.tex
paper

\section*{References}
\begin{thebibliography}{20}
\bibitem{Bell1978} Bell A R 1978 \textit{MNRAS} \textbf{182} 147
\bibitem{Blandford1978} Blandford R D and Ostriker J P 1978 \textit{ApJ} \textbf{221} L29 
\bibitem{Bykov2014} Bykov A M, Ellison D C, Osipov S M and Vladimirov A E 2014 \textit{ApJ} \textbf{789} Issue 2, 137
\bibitem{Berezhko2003} Berezhko E G, Ksenofontov L T and V{\"o}lk H J  2003 \textit{A}{\&}\textit{A} \textbf{412} L11
\bibitem{Uchiyama2007} Uchiyama Y, Aharonian F A, Tanaka T, Takahashi T and Maeda Y 2007 \textit{Nature} \textbf{449} 576
\bibitem{Romansky2016} Romansky V I, Bykov A M, Osipov S M and Gladilin P E \textit{Journal of Physics: CS}
\bibitem{Lapenta2006} Lapenta G, Brackbill J U and Ricci P 2006 \textit{Phys. Plasmas} \textbf{13} 055904
\bibitem{Noguchi2007} Noguchi K, Tronci C, Zuccaro G and Lapenta G 2007 \textit{Phys. Plasmas} \textbf{14} 042308
\bibitem{Ellison2013} Ellison D C, Warren D C and Bykov A M 2013 \textit{ApJ} \textbf{776} Issue 1, 46
\bibitem{Sironi2011} Sironi L and Spitkovsky A 2011 \textit{ApJ} \textbf{741} Issue 1, 39
\end{thebibliography}