### 2017-11-14 papers/Physica2017/intro.tex
paper

\section{Introduction}
One of the possible sources of high-energy cosmic rays are shock waves in astrophysical collisionless plasma due to the diffusive shock acceleration (DSA) mechanism, see \cite{Bell1978}, \cite{Blandford1978}. The particle acceleration on the shock front is sophisticated and nonlinear problem, because cosmic rays have strongly influence on the shock wave. The pressure of cosmic rays is comparable with pressure of background plasma and significatly modifies the front (see for example \cite{Bykov2014}). Also there are a instabilities in anisotropic plasma, which lead to strong field amplification and the observations of the synchrotron emission from supernova remnants (SNRs) show, that the magnetic field there is about 100 times stronger then the interstellar field (see for example \cite{Berezhko2003},\cite{Uchiyama2007}). One of the most powerful methods to explore processes in collisionless plasma is the Particle-in-Cell simulation.