### 2017-11-14 papers/Physica2017/shock.tex

\section{Shock wave simulation}
We studied the evolution of trans-relativistic shocks and spectum of accelerated particles. We chose relastivistic shocks with low lorenz-factor because the maximum energy of produced cosmic rays increases with shock wave lorenz-factor, but the efficiency of acceleration decreases, because it is more difficult for particle to cross fast moving front many times \cite{Ellison2013}. So we assume that intermediate case of trans-relativistic shocks provides the most effective acceleration.

We developed the implicit particle-in-cell (PIC) code, based on the scheme suggested by Lapenta et al.~\cite{Lapenta2006} and improved for the relativistic case by Noguchi et al.\cite{Noguchi2007}.
Our code is fully three dimensional and parallelized with MPI technology, which is adapted for distributed computing and can be executed on a wide class of computers.

In simulation setup the homogeneous plasma, which flows in through the right boundary
and collides with the reflecting superconducting wall on the left boundary.

 Simulations are one-dimensional and have following parameters: initial flow lorenz-factor $\gamma = 1.5$, number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ , temperature $5\cdot10^8 \rm{K}$, magnetic field $B = 10^{-4} \rm{G}$, the full size of the box $L = 4\cdot10^{12} \rm{cm}$, the number of cells $N=4\cdot10^4$. Electron mass is reduced to $m_e = \frac{m_p}{20}$. The full time of simulation is $T = 5000 {\omega_p}^{-1}$. $\theta$ is angle between flow velocity and magnetic field. We present results of particle spectrum in several simulations with different $\theta$. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{fig/protons.eps} 
	\caption{Distribution of protons.}
	\label{protons}
\end{figure}
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{fig/electrons.eps} 
	\caption{Distribution of electrons.}
	\label{electrons}
\end{figure}

The results presented in Figures \ref{protons}-\ref{electrons_with_alpha} show that the values of $F(p)p^4$ for high energies are much greater for protons, than for electrons in both simulations. Also electrons need much more time to be accelerated and to form spectrum $\propto p^{-4}$. It means that protons are injected into the acceleration process more efficiently, and it is consistent with the work of Park et al. {\cite{Park2015}}. Also we have shown that minority of heavy ions increases the spectrum of protons and do not have influence on the spectrum of electrons. It can be explained by the fact, that heavy ions form the large scale turbulence and protons can efficiently scatter on this turbulence. Otherwise for electrons turbulence produced by protons is already enough large-scale  and the influence of heavy ions is neglectable.