### 2017-11-17 papers/Physica2017/conclusions.tex

\section{Conclusions}
In this paper we studied trans-relativistic shock wave using our implicit particle-in-cell code. Simulations show, that particles injection in acceleration process highly depends on angle between flow bulk velocity and magnetic field. Also injection of protons is more effective than electrons for angles less then critical, but there is regime of more effective electrons acceleration when angle is close to critical. Further research is necessary in order to study this phenomena in detail.
\ack
V I Romansky acknowledges support from RSF grant 16-12-10519.
The results of the work were obtained using computational resources of Peter the Great Saint-Petersburg Polytechnic University Supercomputing Center (http://www.spbstu.ru).