### 2017-11-17 papers/Physica2017/Romansky_Physica2017.tex
paper

\documentclass[a4paper]{jpconf}
\bibliographystyle{iopart-num}
\usepackage{amsmath}
\usepackage{citesort}
\usepackage{subfigure}
\usepackage{graphicx}
\graphicspath{{fig/}}
\usepackage{ifpdf}
\ifpdf\usepackage{epstopdf}\fi
\usepackage[export]{adjustbox}

%----------------------------------------------------- 
%\usepackage{soul,ulem,color,xspace,bm}
% Andrei's commands
% Suggest to remove
%\newcommand{\asrm}[1]{{\color{magenta}\sout{#1}}}
% Suggest to insert
%\newcommand{\as}[1]{\color{cyan}#1\xspace\color{black}}
% Suggest to replace
%\newcommand{\asrp}[2]{\asrm{#1} \as{#2}}
% Comment
%\newcommand{\ascm}[1]{{\color{green}\;AS: #1}}
%------------------------------------------------------

\begin{document}
\title{Electron and ion acceleration by relativistic shocks: particle-in-cell model simulations}

\author{V I Romansky$^{1,2}$, A M Bykov$^{1,3}$ and S M Osipov$^1$}

\address{$^1$ Ioffe Institute, 26 Politekhnicheskaya st., St. Petersburg 194021, Russia}
\address{$^2$
	Sternberg Astronomical Institute, Moscow State University
	Universitetsky pr., 13, Moscow 119234, Russia}
\address{$^3$ Peter the Great St.~Petersburg Polytechnic University, 29 Politekhnicheskaya st., St. Petersburg 195251, Russia}

\ead{romanskyvadim@gmail.com}

\begin{abstract}
	The problem of particle acceleration by the collisionless astrophysical shocks is of fundamental importance for the cosmic ray origin problem and the high energy astrophysics. The fast magnetohydrodynamical shocks produced by relativistic jets are considered as the very likely sites of the cosmic ray acceleration.
	While cosmic ray acceleration by both non-relativistic and ultra relativistic shocks was studied in some details the shocks of a moderate Lorentz factors has attracted much less attention so far, while they are important for modelling the gamma-ray burst afterglows and a number of other interesting objects. In this paper we present the results of simulations of relativistic shock waves in the electron-proton plasmas obtained with an implicit particle-in-cell code for the initial study the particle injection efficiencies.
\end{abstract}
\input{intro.tex}
\input{shock.tex}
\input{conclusions.tex}
\input{biblio.tex}
\end{document}