### 2017-11-17 papers/Physica2017/conclusions.tex

\section{Conclusions}
In this paper we presented the first results of the PIC simulations of particle acceleration by the moderately relativistic shock waves. Simulations show, that the injection of particles into the acceleration process depends strongly on the angle between the flow bulk velocity and the magnetic field. The injection efficiency as a function of the inclination angle is studied. The injection of protons is more efficient than that of electrons for the shock inclination angles below the critical angle where the transition to the superluminal shock regime occurs. On the other hand the electron acceleration is more efficient in the shocks with the inclination angles which are close to the critical angle.
\ack
V I Romansky acknowledges support from RSF grant 16-12-10519.
The results of the work were obtained using computational resources of Peter the Great Saint-Petersburg Polytechnic University Supercomputing Center (http://www.scc.spbstu.ru).