### 2017-11-17 papers/Physica2017/intro.tex

\section{Introduction}
One of the possible sources of the high-energy cosmic rays are the shock waves in astrophysical collisionless plasma due to the diffusive shock acceleration (DSA) mechanism, see \cite{Bell1978}, \cite{Blandford1978}. The particle acceleration on the shock front is a sophisticated and nonlinear problem, because the cosmic rays have a strong influence on the shock wave. The pressure of cosmic rays is comparable with the pressure of background plasma and significantly modifies the front (see, e.g., \cite{Bykov2014}). 
Also the instabilities in the anisotropic plasma can yield a strong field amplification, and according to the observations of synchrotron emission from supernova remnants (SNRs), the magnetic field there is about 100 times stronger than the interstellar field (see, e.g., \cite{Berezhko2003},\cite{Uchiyama2007}). 
One of the most powerful methods to explore the processes in collisionless plasma is the Particle-in-Cell simulation.