### 2021-04-14 papers/Physica2021/Romansky_Physica2021.tex
paper

\documentclass[a4paper]{jpconf}
% \bibliographystyle{iopart-num}
\usepackage{amsmath}
\usepackage{citesort}
\usepackage{subfigure}
\usepackage{graphicx}
\graphicspath{{fig/}}
\usepackage{ifpdf}
\ifpdf\usepackage{epstopdf}\fi
\usepackage[export]{adjustbox}

%----------------------------------------------------- 
%\usepackage{soul,ulem,color,xspace,bm}
% Suggest to remove
%\newcommand{\asrm}[1]{{\color{magenta}\sout{#1}}}
% Suggest to insert
%\newcommand{\as}[1]{\color{cyan}#1\xspace\color{black}}
% Suggest to replace
%\newcommand{\asrp}[2]{\asrm{#1} \as{#2}}
% Comment
%\newcommand{\ascm}[1]{{\color{green}\;AS: #1}}
%------------------------------------------------------

\def\apj{ApJ}
\def\mnras{MNRAS}
\def\nat{Nat}
\def\prd{Phys. Rev. D}
\def\araa{ARA\&A}                % "Ann. Rev. Astron. Astrophys."
\def\aap{A\&A}                   % "Astron. Astrophys."
\def\aaps{A\&AS}                 % "Astron. Astrophys. Suppl. Ser."
\def\aj{AJ}                      % "Astron. J."
\def\apjs{ApJS}                  % "Astrophys. J. Suppl. Ser."
\def\pasp{PASP}                  % "Publ. Astron. Soc. Pac."
\def\apjl{ApJ}                   % letter at ApJ
\def\pasj{PASJ}
\def\apss{Astroph. Space Sci.}
\def\aplett{Astroph. Lett}
\def\ssr{Space Sci. Rev.}
\def\aapr{Astron. Astroph. Reviews}
\def\physrep{Phys. Reports}
\def\memsai{Mem. Societa Astronom. Italiana}
\def\jgr{JGR}
\def\jcap{Journal of Cosmology and Astroparticle Physics}

\usepackage{xspace}
\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
% Suggest to remove
\newcommand{\asrm}[1]{{\color{OrangeRed}\sout{#1}}}
% Suggest to insert
\newcommand{\as}[1]{\color{RoyalBlue}#1\xspace\color{black}}
% Suggest to replace
\newcommand{\asrp}[2]{\asrm{#1} \as{#2}}
% Comment
\newcommand{\ascm}[1]{{\color{ForestGreen}#1}}

\begin{document}
	\title{PIC simulation of instabilities in relativistic supernovae shocks}

	\author{V I Romansky$^{1}$, A M Bykov$^{1,2}$ and S M Osipov$^{1}$}

	\address{$^1$ Ioffe Institute, 26 Politekhnicheskaya st., St. Petersburg 194021, Russia}
	\address{$^2$ Peter the Great St. Petersburg Polytechnic University, 29 Politekhnicheskaya st., St. Petersburg 195251, Russia}

	\ead{romanskyvadim@gmail.com}

	\begin{abstract}
                 Modern observations show presence of relativistic supernovae - objects intermediate between regular supernovae and gamma-ray bursts. The typical lorentz-factors of plasma flows in relativistic supernovae are about $1.5$. The process of cosmic ray acceleration is very different from both non-relativistic and ultra-relativistic case. In this paper is presented particle-in-cell modeling of trans-relativistic collisionless shocks. It is shown, that development of magneto-hydrodynamic instabilities ii electron-ion relativistic shock, particularly Bell's instability, has strong influence on cosmic ray acceleration. At times larger than $1500$ reverse ion plasma frequency magnetic field, generated by Bell's instability, significantly suppresses the electrons spectrum. This effect should be taken into account for modeling of synchrothron spectrum from relativistic supernovae.
	\end{abstract}

	\section{Introduction}
	Relativistic shocks are widely discussed phenomena in modern astrophysicist. They can be generated in the supernovae \cite{2010Natur.463..513S,2007ApJ...667..351W}, the accreting supermassive black holes \cite{1984RvMP...56..255B}, the stellar masses black holes \cite{2019MmSAI..90...57M,1999PhR...314..575P,2014LNP...876.....R} and pulsar wind nebulae \cite{2019MNRAS.488.5690O,2017SSRv..207..235B,2017JPlPh..83e6301K,2019ApJ...876L...8B}. Relativistic shocks in astrophysical objects are sources of cosmic rays \cite{2012SSRv..173..309B,2015SSRv..191..519S,2017SSRv..207..319P} and can accelerate particles to ultra-high energies \cite{2009JCAP...11..009L}.

	The presence of high energy non-thermal particles leads to development of various magneto-hydrodynamic instabilities. Bell instability \cite{Bell04} can be the most significant due to it's large growth rate. Instabilities modify the flow and has strong influence on process of particle acceleration. Below we present results of particle-in-cell modeling of relativistic shock and show that growth of magnetic field due to instabilities decreases efficiency of electron acceleration on shock front.

	\section{Numerical setup}
	In this work we use particle-in-cell code Smilei \cite{Smilei18} for modeling trans-relativistic collisionless shocks. We initialize two-dimensional space domain, with electron-proton plasma flowing into simulation box through the right boundary and reflecting wall on the left boundary. In the transversal direction boundary conditions are periodic.

	The simulation parameters are : the initial flow Lorentz factor $\Gamma = 1.5$, the flow magnetization $\sigma = \frac{B^2}{4\pi\Gamma (n_p m_p + n_e m_e) c^2} = 0.004$. The dimensionless thermal energy $\Delta \gamma = \frac{k T}{m_p c^2}$ is equal to $10^{-4}$ and the electron mass is increased up to $m_e = \frac{m_p}{100}$. The size of the simulation box along the $x$ axis is $L_x = 30000\frac{c}{\omega_p}$ and in the transverse direction $L_y = 200\frac{c}{\omega_p}$, where $\omega_p = \sqrt{\frac{4\pi q^2 n}{\Gamma m_e}}$ is the plasma frequency. These scales correspond to $150000$ and $1000$ grid points in $x$ and $y$ directions, respectively.

	Inclination angle of magnetic field due to shock velocity is $\theta = 30$ degrees, which corresponds to the maximum efficiency of particle acceleration \cite{Romansky18} and magnetic field is lying in the simulation plane.

	\section{Results}

	Profiles of transversal magnetic field, obtained from numerical simulation, show significant field amplification in precursor of shock wave, see figure \ref{field}. It is result of development of instabilities in plasma, particularly Bell instability \cite{Bell04}, which growth rate is larger then growth rate of other instabilities.

	\begin{figure}[h!]
		\centering
		\includegraphics[width=0.8\textwidth]{fig/Bnorm.png} 
		\caption{Evolution of the magnetic field in time.}
		\label{field}
	\end{figure}

	Spectrum of electrons is shown in figure \ref{spectrume}. One can see, that electrons distribution function at high energies $(\gamma > 300)$ is maximal at times about 1500 inverse protons plasma frequencies and at the longer times spectrum decreases. It may by explained as consequence of magnetic field amplification. Instabilities increase the transverse magnetic field and also the angle between field and shock velocity. As a result shock becomes quasi-perpendicular and can not effectively accelerate particles \cite{Sironi2011}. On the other hand, protons distribution function constantly increases in time, as shown in figure \ref{spectrump}. We assume, that oscillating transverse field has the strongest influence on the smallest scales, and suppresses he electrons injection in acceleration process.

	\begin{figure}[h!]
		\centering
		\begin{minipage}{0.8\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrum.png} 
			\caption{Temporal evolution of electrons spectrum.}
			\label{spectrume}
		\end{minipage}\hfill
		\begin{minipage}{0.8\textwidth}
			\centering
			\includegraphics[width=0.98\textwidth]{fig/spectrump.png} 
			\caption{Temporal evolution of protons spectrum.}
			\label{spectrump}
		\end{minipage}
	\end{figure}

	\section{Conclusions}

	Particle-in-cell simulations have been used to derive effective macroscopic parameters of relativistic shocks in space plasmas. The effective adiabatic index and electron temperatures are two important parameters required to supplement large-scale single-fluid magneto-hydrodynamic simulations of processes in astrophysical plasmas. The adiabatic indexes for the downstream flow derived with PiC method are summarized in Table 1 and can be used for MHD modeling of relativistic
	supernovae. Further modelling is needed to study the dependence of these two parameters on the initial conditions which, reflect the presence of MHD large-scale turbulence in the stellar wind of the progenitor massive star upstream of the  mildly-relativistic shock of the supernova.

	\ack
	V I Romansky and A M Bykov acknowledge a support from RSF grant 16-12-10225.
	Results of the work were obtained using computational resources of Peter the Great Saint-Petersburg Polytechnic University Supercomputing Center (http://scc.spbstu.ru)

	\section*{References}
	% \bibliographystyle{apalike}
	\bibliographystyle{iopart-num}

	\bibliography{bibliogr}
\end{document}