### 2023-12-12 papers/answer to reviewer.txt
fz

Thanks to the reviewers for important  remarks.

reviewer 1
1) line 35: Nonthermal X-ray luminosity is 10^39 erg/s as written later in the paper.
mistake is corrected
2) Fig.5  : it should be E^2F(E) on the vertical axis
F(E) is defined as energy flux spectrum density, not number of photons spectrum density, so no, in Fig. 4 it should be EF(E)

reviewer 2
1. It is not clear how the authors treat magnetic field amplification—which has bearing on the maximum energy and synchrotron luminosity. How does magnetic field amplification proceed in the Monte Carlo model? In principle, I would expect the non-resonant hybrid instability (Bell 2004) to dominate in this environment. In that case, the amplified field will depend on the acceleration efficiency (and thus the shock inclination) as well as the shock velocity. Regardless, a discussion of magnetic field amplification and its dependence on parameter choices is warranted here.
2. The authors obtain a best-fit shock velocity that is substantially faster than that reported in the literature (Coppejans et al., 2020) despite surprisingly similar inferred densities and magnetic fields. The authors should explain the origins of this discrepancy.
3. CSS161010 was observed at multiple epochs (Coppejans et al., 2020, see their Figure 4). While the most extensive multi-wavelenth SEDs were obtained at t=98 days post-explosion (the epoch considered in the manuscript), there are also detailed radio SEDs at t=68, 162, and 357 days. It would be useful to test whether the authors' model can fit the radio data at these epochs, as well as whether the inferred shock evolution is reasonable. In particular, are these observations consistent with the authors' assumption that CSS161010 does not slow down until at least a year after explosion (line 160)?
4. Related to Item 3: the radio SEDs shown in Coppejans et al., 2020 are far more detailed (>>4 data points) than what is shown in the manuscript. With the additional data points, does model B30 still produce a good fit?
5. While the authors argue that their calculated X-ray luminosity is consistent with observations, the normalization of the SED shown in their Figure 5 looks suspiciously small. I would like to see the X-ray and radio data over-plotted.
6. Section 5 (the jet-massive star interaction model) is interesting, but underdeveloped. I think it is important to make the point that the X-rays may come from another source (especially if, as I suspect, the shock-powered X-rays are a bit low). However, some specifics about the numerical simulations conducted are warranted (e.g., what code was used?). I would also like to see the resulting X-ray SED with data overplotted (both for the CSS161010 model and for the binary, AT2018-like scenario).

Minor points:
1. Given that the authors assumed a free-escape boundary that is half the size of the system, the 100 PeV maximum energy quoted in the abstract is a very optimistic upper limit and should be presented as such.
2. Please elaborate on the assumptions and limitations of the PIC simulations performed in Section 2 (e.g., how many dimensions were used? How realistic was the electron to proton mass ratio?).
3. When referring to the relationship between shock obliquity and acceleration efficiency, the authors should also cite Caprioli and Spitkovsky 2014.
4. Please specify the units of the y-axes in Figures 1-3, preferably in a manner that allows the reader to approximate an acceleration efficiency. Please also specify the obliquity considered in Figure 3.
5. In Figures 1-2, it might be useful to plot in terms of particle momentum (p^4f(E) vs p) order to make the supra-thermal bump more apparent and to eliminate features due to the transition from non-relativistic to relativistic protons.
6. Please clarify how the radio SED prediction would change if a sphere were assumed instead of cylinder (i.e., if there were regions of the source that were optically thicker/thinner at a given frequency)
7. It would be helpful to clarify early on that the magnetic field used to compute the synchrotron emission in Figures 4 and 5 comes directly from Monte Carlo simulations. Given assumptions of density, shock velocity, and magnetic inclination, acceleration efficiency and thus the amplified magnetic field are derived parameters. As such, the F(nu, B, n) notation starting around line 231 is confusing since, as far as I can tell, B is not a model input.
8. It would also be helpful to identify, either in Figure 5 or in the text, key features in the multiwavelenth SED. As I understand it, the first peak is nu_SSA, the dip corresponds to the hardening of the electron spectrum due to non-linear DSA, and the second peak is the cooling break.
9. Regarding the discussion of the strong inferred magnetic field starting around line 252: perhaps the origin of this magnetic field is simply CR-driven magnetic field amplification which, as I understand it, is already included in the Monte Carlo model?